\documentclass{amsart}

\input{preambleGen}

\title{Coproduct of Operads}
\author{Daniel R. Barrero R.}
\date{\today}

\begin{document}

% -	C.f. work log of June 28	- %

\maketitle

\section{Trees}

\input{trees}

\section{Free operads}

Let $U : \mathrm{Op} \to \mathrm{Top}^{\NN}$ be the forgetful functor and
let $F$ be its left adjoint. Its existence is a result of Boardman and Vogt
in \cite{bv-hiasots}.

\begin{defn}
	Let $C$ be a collection. A \emph{$C-$labelled planar $n-$tree} is a
	planar $n-$tree such that each vertex with $k$ children is labelled
	by an element of $C_k$.
\end{defn}


	

\section{The construction}

Given two operads $O$ and $O'$, the $F-U$ adjoint pair gives epimorphisms 

\begin{eqnarray}\label{fu-epis}
	\epsilon : FU \ O \to O \\
	\epsilon' : FU \ O' \to O'.
\end{eqnarray}

If the category $\mathrm{Op}$ had coproducts, we would have an operad
epimorphism\footnote{It being an epimorphism is not immediate, but the
property is necessary for the validity of the construction.}

\begin{equation}\label{cpd-epis}
	\epsilon + \epsilon' : (FU \ O) + (FU \ O') \to O + O'.
\end{equation}

Drawing inspiration\footnote{We haven't established nor cited any relevant
categorical properties for operads.} from the fact that left adjoints
preserve colimits \cite{riehl-ctic}, we may define


\begin{equation}\label{fu-coprod}
	(FU \ O) + (FU \ O') \ := \ F \ \left( (U \ O) + (U \ O')
	\right).
\end{equation}

and therefore compute a quotient of $F \ \left( (U \ O) + (U \ O')
\right)$ that is a coproduct of $O$ and $O'$.

\section{Algebras}


In order for $A$ to be an algebra over $O + O'$, it only requires having
algebra structrues over both $O$ and $O'$, with no compatibility conditions
imposed:

\begin{equation}\label{opd-cpd}
		\begin{tikzcd}
			& O + O' \arrow[dd, dashed, "u + v" description] &
					\\
			O \arrow[rd, "u"'] \arrow[ru, "i"] & & O'
				\arrow[ld, "v"] \arrow[lu, "j"'] \\
			& \mathrm{End}(A) &
		\end{tikzcd}
\end{equation}

\section{Comments}

\subsection{} The result by B \& V is that the forgetful functor is
\emph{monadic.}

\subsection{} The full statement that motivates \eqref{fu-coprod} is that
\emph{Right adjoints preserve limits and left adjoints preserve colimits.}
It can be found in \cite{riehl-ctic}.

\subsection{} The definition \eqref{fu-coprod} suggests that
\emph{``at the $FU-$level coproducts are trivially definied.''}

\subsection{} Justify that $\epsilon + \epsilon'$ in \eqref{cpd-epis} is
an epimorphism.

\subsection{} The number of edges of a tree given by definition
\ref{def-tree} is probably

$$
(\# \ V) + n,
$$

where $n$ is the number of leaves.

\subsection{Yoneda's lemma} I must write down this simple but important
fact at some point.

\subsection{General products and coproducts}

\begin{defn}\label{prod-def}
	Let $X$ and $Y$ be objects of the category $\mathcal{C}$. Their
	\emph{product}, if it exists, is the data $(X \times Y, p, q)$
	uniquely determined by the property that for any object $Z$ of
	$\mathcal{C}$ and maps $f : Z \to X$ and $g : Z \to Y$, the
	following diagram commutes:
	\begin{equation}\label{prod-cd}
		\begin{tikzcd}
			& X \times Y \arrow[ld, "p"'] \arrow[rd, "q"] & \\
			  X &            & Y \\
			  & Z \arrow[lu, "f"] \arrow[ru, "g"']
			      \arrow[uu, dashed, "f \times g" description] & 
		\end{tikzcd}
	\end{equation}
\end{defn}

We obtain the definition of coproduct by inverting the directions of the
arrows in definition \ref{prod-def}:

\begin{defn}\label{coprod-def}
	The \emph{coproduct} of $X$ and $Y$, if it exists, is the data
	$(X + Y, i, j)$ uniquely determined by the property that for any
	object $Z$ and maps $u : X \to Z$ and $v : Y \to Z$, the following
	diagram commutes:
	\begin{equation}\label{coprod-cd}
		\begin{tikzcd}
			& X + Y \arrow[dd, dashed, "u + v" description] &
					\\
			X \arrow[rd, "u"'] \arrow[ru, "i"] & & Y
				\arrow[ld, "v"] \arrow[lu, "j"'] \\
			& Z &
		\end{tikzcd}
	\end{equation}
\end{defn}

\bibliographystyle{acm}
\bibliography{refs}

\end{document}
